### 2018-10-18 Thesis/Background.tex

\chapter{A Dynamical Recipe for Cosmological Disks}\label{ch:Background}

\section{Physical Motivation of Modeling}

\section{Time Evolution of Collisionless Systems}
\subsection{The Collisionless Boltzmann Equation}
\subsection{Numerical Solutions}

\section{Phase Space, Equilibrium, and Initial Conditions}
\subsection{Jeans Modeling}
\subsection{DF-based Models and the Strong Jeans Theorem}
\subsection{Action-Angle Variables}
\subsection{DFs as Functions of Actions}

\section{Simulation Analysis: Paradigms and Tools}
\subsection{WKB Wave Analysis}
\subsection{Time Series Filtering}
\subsection{MCMC}

\section{Cosmology and Implications for Galaxy Studies}
\subsection{Basic FRW Cosmology}
\subsection{Extension of Numerical Methods}
\subsection{Sampling Initial Conditions for Cosmological Simulations}
\subsection{Cosmological Substructure and Halo Triaxiality}
\subsection{Identifying Substructure in Simulations}