### 2018-10-18 Thesis/Thesis.tex

%*************************************************************************************************************
% PREAMBLE STUFF
%*************************************************************************************************************
% Instead of inserting my \usepackage and defined commands here, I keep them in a separate file
\include{myPreamble}

%*************************************************************************************************************
% INCLUDE ONLY
%*************************************************************************************************************
% Use if you want to include only certain parts of the document, example \includeonly{introduction}
% in order to speed up compile time when you're focussing on some particular part.
%\includeonly{}

%*************************************************************************************************************
% DOCUMENT
%*************************************************************************************************************
\newcommand{\zrms}{z_{\rm rms}}
\newcommand{\zeff}{z_{\rm rms,eff}}
\renewcommand{\thepage}{}
%\newcommand{\last_page_of_biblio}{\def \last_page_number \numpages}
%\newcommand{\stitch_to_last_chapter}{\setcounter{page}{\last_page_number}}

\begin{document}

%*************************************************************************************************************
% TITLE
%*************************************************************************************************************

\title{Nature or Nurture? Collisionless Evolution of Galactic Disc-Halo Systems}

\author{J S Bauer}

\dept{Department of Physics}
\degree{Doctor of Philosophy}

% OPTIONAL HERE
% ~~~~~~~~~~~~~
% \submitdate{month year in which submitted to GPO}
%        - date LaTeX'd if omitted
% \copyrightyear{ear degree conferred}
%        - year LaTeX'd if omitted
% \figurespagetrue or \figurespagefalse
%        - produce or don't produce a List of Figures page (true by default)
% \tablespagetrue or \tablespagefalse
%        - produce or don't produce a List of Tables page (true by default)

\beforepreface

% Adding single spacing so abstract and table of contents is single spaced.

%*************************************************************************************************************
% ABSTRACT
%*************************************************************************************************************

\prefacesection{Abstract}

\input{Abstract}

%*************************************************************************************************************
% CO-AUTHORSHIP (if necessary)
%*************************************************************************************************************

%\prefacesection{Co-Authorship}

%*************************************************************************************************************
% ACKNOWLEDGEMENTS
%*************************************************************************************************************

\prefacesection{Acknowledgments}

Blah blah blah.

%*************************************************************************************************************
% STATEMENT OF ORIGINALITY (required CHEM, CISC, GEOL, MATH, PHYS (Ph.D. only))
%*************************************************************************************************************

\prefacesection{Statement of Originality}

Only required by CHEM, COMPUTING, GEOL, MATH and Physics (Ph.D. ONLY!).

\singlespacing \afterpreface \doublespacing

%*************************************************************************************************************
% MEATY CHAPTERS
%*************************************************************************************************************

%Testing my index\index{index} abilities.  And
%sub-entry\index{index!subentry} abilities.

% This command can be used to view the page layout for this document
% \layout

% Here, I'm tweaking how much space is put above and below floats.
% Comment out if you want the *purest* latex spacing.
%\setlength{\abovedisplayskip}{3pt plus1pt minus1pt}
%\setlength{\abovedisplayshortskip}{3pt plus1pt minus1pt}
%\setlength{\belowdisplayskip}{3pt plus1pt minus1pt}
%\setlength{\belowdisplayshortskip}{3pt plus1pt minus1pt}

% Include my chapter texts - kept separated to make editing easier.
%\include{Glossary}
\include{Introduction}
\include{Background}
\include{Paper_I}
\include{Paper_II}
%\include{Alloy}

%\include{Embee_UserPerspective}
%\include{Embee_Implementation}

\include{Conclusion}
%*************************************************************************************************************
% BIBLIOGRAPHY
%*************************************************************************************************************
% This GATHER command is useful for when you want to use WinEdt's Gather functionality, i.e., type
% \cite{} and a popup box appears with all of your citations to choose from.  Leave the % on the next line.
%GATHER{thesis.bib}

% Put in \nocite{*} so all entries in the bibliography are included
%\nocite{*}

%\bibliographystyle{apalike}
%\bibliography{../bibliography.bib}
\include{Appendices}

%*************************************************************************************************************
% APPENDICES
%*************************************************************************************************************
\updatechaptername
%\appendixpage
\appendix

%\include{App_AnalyzerCode}
%\include{App_ExtraAnalysis}
%\addappheadtotoc

%*************************************************************************************************************
% GLOSSARY
% Using a glossary is more than beginners need to know; leaving the packages, etc. here for now.
%*************************************************************************************************************
%\usepackage[nonumberlist]{glossaries}
%\usepackage[refpages]{gloss}  % for my glossary
                              % refpages shows the first page where the term occurs
%-------------------------------------------------------------------------------------------------------------
% Tell Latex to make a glossary
%*************************************************************************************************************
%\makeglossaries  % tell latex to make the glossary
%\glossarystyle{list}

%*************************************************************************************************************
% INDEX
%*************************************************************************************************************
% Here's where the index would be printed, if you created one.  Remove the % on the next line.
%\printindex

%*************************************************************************************************************
\end{document}