### 2019-07-15 Thesis/Paper_III.tex

\chapter{Forming Vertical Structure in $\Lambda$CDM Discs}\label{ch:paper_iii}