### 2019-07-16 Thesis/Appendices.tex
Semi-final presubmission draft

\appendix

\definecolor{codegreen}{rgb}{0,0.6,0}
\definecolor{codegray}{rgb}{0.5,0.5,0.5}
\definecolor{codepurple}{rgb}{0.58,0,0.82}
\definecolor{backcolour}{rgb}{0.95,0.95,0.92}

\chapter{Referenced Code: kicks.c} \label{ch:kicks.c}
\lstinputlisting[language=C,basicstyle=\footnotesize,
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},]{kicks.c}]

\chapter{Referenced Code: predict.c} \label{ch:predict.c}
\lstinputlisting[language=C,basicstyle=\footnotesize,
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple}]{predict.c}]

\chapter{Referenced Code: extract\_halo\_ascii.py} \label{ch:extract_halo_ascii.py}
\lstinputlisting[language=Python,basicstyle=\footnotesize,
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},]{extract_halo_ascii.py}]

\chapter{Referenced Code: merge\_ics.cpp} \label{ch:merge_ics.cpp}
\lstinputlisting[language=C++,basicstyle=\footnotesize,
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},]{merge_ics.cpp}]

\chapter{Referenced Code: timestep.c} \label{ch:timestep.c}
\lstinputlisting[language=C,basicstyle=\footnotesize,
    commentstyle=\color{codegreen},
    keywordstyle=\color{magenta},
    numberstyle=\tiny\color{codegray},
    stringstyle=\color{codepurple},]{timestep.c}]